Proposed Experimentation

### 2017-04-06 reports/experiment.tex
Update experiment plan

@@

The experiments must be designed to answer the important questions about the
detour routing schemes. In particular:

\begin{itemize}
\item Does detour routing improve connection performance when the main route is
  unreliable? How much does it improve performance, and by what performance
  metrics?
\item Is there a measurable performance difference between the OpenVPN and NAT
  approaches to detour routing?
\item Is there a measurable performance difference between scheduling modules in
  this scenario?
\end{itemize}

To quantify performance, the following metrics will be useful:
\begin{itemize}
\item Throughput - total bytes transferred through a connection per unit time.
\end{itemize}

\subsection{Small Scale}

First, we should validate that the framework performs as expected on a
controlled network topology. A reasonable test framework would then be to set up
a network topology with multiple routes between a client and server, vary these
parameters, and compare the performance metrics to those of several controls:

\begin{itemize}
\item Vanilla TCP, using the route which has the ``lossy'' link.
\item Vanilla TCP, using the alternative route through the detour.
\item Multipath TCP, unmodified. In this scenario, the network topology would be
  altered such that the client has two interfaces and each has a different route
  to the server.
\end{itemize}

I would hypothesize that the detour routing solution will outperform the first
option, and it would be comparable to the rest.

% TODO: diagram of proposed topology

\subsection{Large Scale}

My ideas of what would make a suitable larger scale test are not as well-formed.
I believe we need a larger scale test to give some indication that this approach
could work ``in the wild''. For this sort of test, rather than setting up a
controlled, repeatable network topology, we would simply use nodes in separate
sites and perform a similar data transfer. The transfer would have to be
repeated (control and MPTCP) several times in a row to give real statistical
backbone to the experiment.

>>>>

\subsection{Mininet}

Mininet is a single-computer network emulation platform. It allows users to
define arbitrary network topologies, including hosts, links, and switches. The
hosts are simply Linux processes (including potentially child processes) running
in a separate kernel network namespace. All networking occurs on the same
kernel, meaning that the machine running the experiment must have a custom
kernel (rather than simply running a virtual machine with the custom kernel).

It should be possible to run experiments on Mininet - the biggest concern would
be ensuring that the required daemons and kernel modifications don't conflict
with each other, since they would be running on the same kernel with only
network namespaces separating them.

As far as performance, making a Mininet experiment ``network-limited'' requires
that you assign bandwidth to each link that is less than the total your CPU
could handle. On my laptop (within a virtual machine), this maximum is around 24
Gbps. A reasonable setup could assign each link to be 100Mbps, have several
links and routers, and not stress the CPU for processing this.

@@

The point of this thesis is to propose a mechanism for applying detour routing
to any TCP connection. While it is out of scope to attempt to determine an
optimal detour routing scheme, some experimentation is necessary, if for no
other reason than to validate that this mechanism can provide the benefits of
Multipath TCP to alternative ``detour'' routes. The main benefit of MPTCP is
that it provides link aggregation, or in this case, path aggregation. That is,
we are able to use the capacity of more than one path in our connection.

Rather than design experiments that attempt to determine the best strategy for
picking a detour, or the best strategy for scheduling data when you have
multiple connections, we should simply show that this implementation provides
these aggregation benefits. There are a couple types of benefits that we should
attempt to quantify:

\begin{itemize}
\item Better throughput. In the case that our connection speed is bottlenecked
  by a link that we can circumvent, we should be able to aggregate bandwidth so
  that we improve throughput.
\item Better response time. In the case that one path has high delay, we should
  be able to improve response time by adding another path.
\item Better fault tolerance. If a path has a high loss rate, we should be able
  to improve our throughput by circumventing that lossy link.
\end{itemize}

So, the experiments are designed around these three scenarios. We use the same
network topologies for each scenario, but we alter the connection properties of
a link along the ``Internet routed path'' (but not the detour routed path).
Then, we compare the desired property (throughput or response time) to the
following controls:

\begin{itemize}
\item Vanilla TCP. No multipath TCP, just a single TCP flow from client to
  server.
\item Vanilla TCP, on the detour route.
\end{itemize}

This way, we can determine whether the MPTCP detour mechanism is able to (at
least) leverage the performance of the better path, if not aggregate their
performance.

These experiments could be run on Mininet or Emulab. In both cases, the results
would be highly reproducible. If Mininet were used, we could run very many
trials very quickly in order to get high statistical power. In addition, for
publication we could provide a complete virtual machine image which could
reproduce all experiments. Finally, if changes were required to the detour
framework (as they will probably be), Mininet will allow for the quickest
response time, so that we can run and re-run our tests quickly.
